---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(ggpubr)
library(reshape2)
library(igraph)
```


Load functions

```{r}
#source("simulation.R")
source("functions_metrics_networks.R")
#source("simulation_nobasals.R")
```

Parameters for plotting

```{r}
my.theme<-theme(axis.text=element_text(size=12),
                axis.title = element_text(size = 14),
                legend.position = "top",
                legend.text=element_text(size=10),
                legend.title = element_text(size=12),
                plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
                axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
                axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)))
```




```{r}
source("simulation_with_competition_fw.R")

```



```{r}

pars = list()

pars$competition_coefficient = 0.01

pars$Sbasal = 25 # number of basal species
pars$Smax = 1000 # Maximal number of species allowed

pars$Bspe = 4 # Basal species impact on the speciation probality
pars$Bext = 4 # Basal species impact on extinction probability

pars$av_r = 0.2 # range of the niche


# PROBABILITY OF MUTATION

pars$u_max = 0.25 #0.15 # mutation probability (0.075, )
pars$d = 0.5 # Decrease speed of the establishment probability
pars$I_max = 60 # Maximal number of interactioning species

pars$beta_n =  1 # parameter of the beta distribution 



# STRENGTH OF NEUTRAL-DRIVEN EVOLUTION

pars$SN = 0 # strength for neutral-driven evolution



# PROBABILITY OF ESTABLISHMENT


# strength of selection-driven selection is 1 - SN
pars$estab_prob_neutral = rep(0, pars$Smax) # neutral probability of establishment

# Competition

#pars$u_0neg = 0.075 # Asymptotic establishment probability with infinite competitive interactions
#pars$u_1neg = 2 # Establishment probability with absence of competitive interactions
#pars$a_uneg = 0.075 # Shape of the exponential decay of the colonization - interaction relationship

# Facilitation & Foodweb

pars$u_0pos = 1  
pars$u_1pos = -1 
pars$a_upos = 0.45 



# PROBABILITY OF EXTINCTION

pars$ext_prob_neutral = 0 # neutral probability of extinction


# Competition

pars$e_0neg = 0.1 #0.15 # Asymptotic extinction probability with infinite negative interactions
pars$a_eneg = 0.025 # Shape of the exponential decay of the negative extinction - interaction relationship
pars$e_1neg = -pars$e_0neg  # Extinction probability with absence of interactions

# Facilitation & Foodweb

pars$e_0pos = 0.075 
pars$e_1pos = 5.19 
pars$a_epos = 1.2 



#saveRDS(pars, "pars.rds")

```


# set parameters simulation

```{r}
nsteps = 250 # Set the maximum timestep per simulation
nsim = 2 # Set the number of simulations
```




### Test one interaction


```{r}

pars$int = 2 # 0 comp, 1 fac, 2 fw

simulation <- sim_model(seed = 100, 
                        pars, 
                        nsteps = nsteps)

res <- list(
                      #seed = seed,
                      parameters = pars, 
                      presence_matrix = simulation$pres, 
                      traits_df = simulation$traits,
                      parentage_matrix = simulation$anc,
                      extinxction_matrix = simulation$extinct, 
                      network_list = simulation$L_list
                      # anc_dist_table = simulation$dist_anc, 
                      # list_anc_dist = simulation$list_dist_anc
      ) # Record results from the simulation





list_results <- list()
list_adj_matrix.f <- list()

for (i in 1:nsim) {
  
  
  seed <- set.seed(50)
  
  simulation <- sim_model(seed = runif(1), 
                        pars, 
                        nsteps = nsteps)
  
  res <- list(
                      #seed = seed,
                      parameters = pars, 
                      presence_matrix = simulation$pres, 
                      traits_df = simulation$traits,
                      parentage_matrix = simulation$anc,
                      extinxction_matrix = simulation$extinct, 
                      network_list = simulation$L_list
                      # anc_dist_table = simulation$dist_anc, 
                      # list_anc_dist = simulation$list_dist_anc
      ) # Record results from the simulation
  
  list_results[[i]] <- res
  
     
  list_adj_matrix.f[[i]] <- get_adjancency_from_simulation.fw_single(list_networks = res$network_list,
                                           nbasals = pars$Sbasal,
                                           pres_mat = res$presence_matrix)
  
}

```






# Explore results


```{r}

# count number of timesteps where there are spp

n_steps <- length(res$network_list)

pres = res$presence_matrix



```



## N values with time

```{r}



res_spp = data.frame(matrix(ncol=2,
                      nrow=0, 
                      dimnames=list(NULL, c("step", "n values")))) #makes an empty dataframe

res_mean = data.frame(matrix(ncol=2,
                      nrow=0, 
                      dimnames=list(NULL, c("step", "n values")))) 


for (i in 1:nsteps) {
  
  pres_stepi <- which(pres[i,] == 1)
  traits_stepi <- res$traits_df[pres_stepi, "n"]
  
  vec_step <- rep(i, times = length(traits_stepi))
  
  
  
  res.add <- data.frame("step" = vec_step, 
                               "n values" = traits_stepi)
  
  res_spp <- rbind(res_spp, res.add)
  

  
  res.add1 <- data.frame("step" = i, 
                                "n values" = mean(traits_stepi))
  
  res_mean <- rbind(res_mean, res.add1)
  
  
}

plot_nichetime <- ggarrange(
  
  ggplot(res_spp, aes(x = step, y = n.values)) +
    geom_point(alpha = 0.2)+
    ggtitle("spp n values")+
    theme_classic()+
    my.theme+
    ylim(0,1),
  
  
  
  ggplot(res_mean, aes(x = step, y = n.values)) +
    geom_point(alpha = 0.2)+
    ggtitle("mean n values")+
    theme_classic()+
    my.theme+
    ylim(0,1),
  
  nrow = 2, 
  ncol = 1
  
  
)

plot_nichetime

```


## final degree distribution

```{r}

net_f <- res$network_list[[length(res$network_list)]]


net_f <- net_f[26:1025,]

colnames(net_f) <- seq(1:1000)

rownames(net_f) <- seq(1:1000)

pres_f <- res$presence_matrix[length(res$network_list),]
names(pres_f) <- seq(1:1000)


net_f <- net_f[names(which(pres_f == 1)), names(which(pres_f == 1))]

graph <- graph_from_adjacency_matrix(adjmatrix = net_f,
                                     mode = "directed")

degree <- degree(graph, mode="all")

deg.dist <- degree_distribution(graph, cumulative=T, mode="all")

spp <- 1:length(deg.dist)

df_degreeplot <- data.frame(deg.dist, spp)

plot_degreedistr <- ggplot(df_degreeplot, aes(x = spp, y = deg.dist))+
  geom_point() +
  theme_classic()+
  my.theme+
  ylab("degree")+
  xlab("spp")

plot_degreedistr

```



## Speces richness with time



```{r}

plot_spp_richness <- function(steps, pres_mat, title) {
  
  timesteps <- 1:steps
  
  n_spp_time <- c()
  
  for (i in 1:nrow(pres_mat)) {
  
  n_spp_time[i] <- length(which(pres_mat[i,] == 1))
  
  }
  
  
  df_spprich_time <- data.frame(timesteps,n_spp_time)
  
  p.spp <- ggplot(df_spprich_time, aes(x=timesteps, y = n_spp_time)) +
  geom_line(color="black", linetype="twodash") +
  theme_classic()+
  my.theme+
  xlab("timesteps")+
  ylab("N species")+
    ggtitle(title)
  
  return(p.spp)
}


plot_richnesstime <- plot_spp_richness(steps = n_steps, pres_mat = res$presence_matrix[1:n_steps,], title = "foodweb")

plot_richnesstime
```

## Speciation and extinction events


```{r}

# Number of speciation and extinction events
t0 = pres[1:(n_steps-1),]
t1 = pres[2:n_steps,] 
spec_mat = pres[1:(n_steps-1),]*0
ext_mat = pres[1:(n_steps-1),]*0
spec_mat[t1-t0==1] = 1 
ext_mat[t1-t0==-1] = 1
spec = apply(spec_mat ,1,sum)
ext = apply(ext_mat, 1, sum)

S = apply(pres,1,sum)[2:n_steps]

time =  1:(n_steps-1)

df_divrates <- data.frame(spec, ext, S, time)

df_plot_divrates <- melt(df_divrates, id = c("time", "S"))


plot_divrates <- ggplot(data=df_plot_divrates, aes(x=time, y=value/S, group=variable)) +
  geom_line(aes(color = variable))+
  scale_color_manual(values = c("blue4", "red2"))+
  theme_classic()+
  my.theme+
  ylab("rate")+
  labs(color = "")

plot_divrates


```



## diversification-richness dependence

```{r}

plot_div_richness <- ggplot(data=df_plot_divrates, aes(x=S, y=value/S, group=variable)) +
  geom_point(aes(color = variable), size = 2, alpha = 0.4)+
  scale_color_manual(values = c("blue4", "red2"))+
  theme_classic()+
  my.theme+
  ylab("rate")+
  xlab("Species richness")+
  labs(color = "")

```



```{r}
arrange_plots <- ggarrange(
  plot_nichetime + ggtitle("evolution niche values"),
  plot_degreedistr + ggtitle("degree distr. final"),
  plot_richnesstime + ggtitle("evolution richness"),
  plot_divrates + ggtitle("evolution div. rates"),
  plot_div_richness +ggtitle("div.rates & richness")
)

arrange_plots



```



## Further checks for foodweb

### Read table metrics from empirical foodwebs

empirical networks data was obtained from the ECOWeB 1.1 database:
https://digitalcommons.rockefeller.edu/cohen_joel_laboratory/1/

```{r}


net_metr_empirical <- readRDS("net_metrics_empirical.rds")

net_metr_empirical$id <- NULL

# eliminate one problematic foodweb:

net_metr_empirical <- net_metr_empirical[-13,]

# reshape dataframe:

df_metrics_emp <- reshape_df_net.metrics(df_metrics = net_metr_empirical,
                       interaction = "empirical")
```



```{r}

list_adj_matrix.f <- list()

  list_adj_matrix.f[[2]] <- get_adjancency_from_simulation.fw_single(list_networks = res$network_list,
                                           nbasals = pars$Sbasal,
                                           pres_mat = res$presence_matrix)
```



```{r}



net_metr_sim <- get_network_measures_from_adjacency(list_adj_matrix = list_adj_matrix.f)

df_metrics_fw <- reshape_df_net.metrics(df_metrics = net_metr_sim,
                       interaction = "foodweb")

df_metrics <- merge(df_metrics_fw,
                    df_metrics_emp,
                    by = "metric")


# plot the results

p_S <- plot_metric_foodweb(df = df_metrics, metric = "S", y_axis = "N species")
p_link.dens <- plot_metric_foodweb(df = df_metrics, metric = "Link_density", y_axis = "L/S")
p_C <- plot_metric_foodweb(df = df_metrics, metric = "C", y_axis = "Connectance")
p_top <- plot_metric_foodweb(df = df_metrics, metric = "perc_tops", y_axis = "% top species")
p_int <- plot_metric_foodweb(df = df_metrics, metric = "perc_int", y_axis = "% intermediate species")
p_basal <- plot_metric_foodweb(df = df_metrics, metric = "perc_basals", y_axis = "% basal species")
p_cannibals <- plot_metric_foodweb(df = df_metrics, metric = "perc_cannibals", y_axis = "% cannibal species")
p_omnivory <- plot_metric_foodweb(df = df_metrics, metric = "omnivory", y_axis = "omnivory")
p_sd_gen <- plot_metric_foodweb(df = df_metrics, metric = "sd_gen", y_axis = "SD generality")
p_sd_vul <- plot_metric_foodweb(df = df_metrics, metric = "sd_vul", y_axis = "SD vulnerability")

plot_metrics_comparison_sim_emp <- ggarrange(
  p_S,
  p_link.dens,
  p_C,
  p_top,
  p_int,
  p_basal,
  p_cannibals,
  p_omnivory,
  p_sd_gen,
  p_sd_vul,
  
  nrow = 5,
  ncol = 2
  
)

plot_metrics_comparison_sim_emp
```






```{r}
annotate_figure(arrange_plots, top = "facilitation")
annotate_figure(arrange_plots, top = "competition")
annotate_figure(arrange_plots, top = "foodweb")



ggsave("fac_res_sim.png", height = 10, width = 13)
ggsave("comp_res_sim.png", height = 10, width = 13)
ggsave("fw_res_sim.png", height = 10, width = 13)
```

