---
title: "Testing parameters for facilitation and competition networks"
---


Script to try parameters out and explore the output.

I haven't found empirical network data for facilitation and competition so far. The evaluation I am doing consist mainly in exploring the species richness curve (expecting a progressive increase with time and then an estabilization), degree distribution (expecting diversity of values), and some model objects (trait evolution, speciation/extinction events).






```{r}
library(EnvStats)
library(foreach)
library(ggplot2)
library(igraph)
library(ggpubr)
library(reshape2)

```


For plotting:

```{r}
my.theme<-theme(axis.text=element_text(size=12),
                axis.title = element_text(size = 14),
                legend.position = "top",
                legend.text=element_text(size=10),
                legend.title = element_text(size=12),
                plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
                axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
                axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)))
```



```{r}
source("../functions/functions_simulation_fac_comp.R")
```



```{r}



pars = list()

# General parameters
#pars$int = 0 # 0 for competition, 1 for facilitation, 2 for predation
pars$Smax = 1000 # Maximal number of species in the system

# Related to interaction range trait
pars$av_r = 0.1 # Half range of the niche of the first species
pars$sd = 0.5*pars$av_r + 0.0001 # Standard deviation of the normal distribution used to calculate the niche optimum trait of a new species

pars$u_max = 0.15 # Speciation probability

# Extinction probability, negative interactions communities
pars$a_eneg = 0.025 # Shape of the exponential decay of the negative extinction - interaction relationship
pars$e_0neg = 0.5 # Asymptotic extinction probability with infinite negative interactions
pars$e_1neg = 1  # Extinction probability with absence of interactions

# Extinction probability, positive interactions communities
pars$a_epos = 1.2  # Shape of the exponential decay of the positive extinction - interaction relationship
pars$e_0pos = 0.075 # Asymptotic extinction probability with infinite positive interactions
pars$e_1pos = 5.19 # 1 - pars$e_0pos

# Establishment probability, negative interactions communities
pars$a_uneg = 0.075 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0neg = 0.075 # Asymptotic establishment probability with infinite competition interactions
pars$u_1neg = 2 # Establishment probability with absence of competition interaction

# Establishment probability, positive interactions communities
pars$a_u = 0.45 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0 = 1 # Asymptotic establishment probability with infinite facilitation interactions
pars$u_1 = -1 # Establishment probability with absence of facilitation interactions
pars$d = 0.5 # Decrease speed of the establishment probability

# Extinction and establishment probabilitie for both, positive and negative interactions communities
pars$Bspe = 4 # Constant minimal number of interaction per species (extablishment prob)
pars$Bext = 4 # Constant minimal number of interaction per species (extinction prob)
pars$I_max = 40 # Maximal number of interactioning species


pars$beta_n = 1 # parameter of the beta distribution


# STRENGTH OF NEUTRAL-DRIVEN EVOLUTION

pars$SN = 0 


# PROBABILITY OF ESTABLISHMENT
# strength of selection-driven selection is 1 - SN
pars$estab_prob_neutral = 0 

# PROBABILITY OF EXTINCTION

pars$ext_prob_neutral = 0 # neutral probability of extinction


#saveRDS(pars, "pars_faccomp_12june.rds")

#pars <- readRDS("pars_work_12june.rds")

```




tests

```{r}
# Extinction probability, negative interactions communities
#pars$a_eneg <- 0.05 # Increased sensitivity
pars$e_0neg <- 0.65  # Increased baseline extinction probability

# Extinction probability, positive interactions communities
#pars$a_epos <- 2.4  # Increased sensitivity
pars$e_0pos <- 0.06 # Decreased baseline extinction probability

# Establishment probability, negative interactions communities
#pars$a_uneg <- 0.15 # Increased sensitivity

# Establishment probability, positive interactions communities
pars$a_u <- 1 # Increased sensitivity

```


```{r}
# Extinction probability parameters
pars$e_0pos <- 0.1  # Reduce the asymptotic extinction probability with infinite positive interactions
pars$e_1pos <- 6  # Slightly decrease the extinction probability in the absence of positive interactions
pars$a_epos <- 0.8  # Decrease the rate at which extinction probability decreases with more positive interactions

# Establishment probability parameters
pars$u_0 <- 1.0  # Increase the asymptotic establishment probability with infinite positive interactions
pars$u_1 <- -0.9 #-0.9  # Increase the establishment probability in the absence of positive interactions
pars$a_u <- 1.0  # Keep the rate at which establishment probability decreases with more positive interactions


```



as Mathilde

- range evolves
- Niche is drawn from a normal distribution

# set parameters simulation

```{r}
nsteps = 250 # Set the maximum timestep per simulation
nsim = 2 # Set the number of simulations

# Vector to record number of simulation needed to reach 100 valid simulations
needed_sim_pos <- c(rep(NA, nsim))
needed_sim_neg <- c(rep(NA, nsim))
```


# Simulation


```{r}

list_results_fac <- list()
list_results_comp <- list()
seed_list <- NULL

	total_nb_sim <- 0
	b <- 1

	for(int in 0:1){
		pars$int = int # Define the interaction type: 0 = competition ; 1 = facilitation
		successful_sim <- 1 # Set the count of simulations
		list_res <- list() # Set the list to store the results

		if(pars$int == 1){
			print("Positive interctions")
		} else {
			print("Negative interctions")
		}

		seed_record <- c(rep(NA, nsim)) # Set the vector to record the seed that have been used

		while(successful_sim < nsim+1){

  		seed <- sample.int(80000, 1) # Pick a random number, this number will be the seed for one simulation
  		print(seed)

			# If a seed have already been tested, pick another one
			if(length(which(seed_list == seed)) != 0){
				while(seed %in% seed_list){
    			seed <- sample.int(80000, 1)
  			}
			}

			seed_list <- c(seed_list, seed) # Add the seed

			# SIMULATION ; from the file code/functions/functions_M-C_bifurcation.R
  		simulation <- sim_model_bif_fc(seed = seed, pars, nsteps = nsteps)  
  	
  		#---------------------- stored: pres, traits_mat, anc, extinct, Stotact, L, L_list

			total_nb_sim <- total_nb_sim +1 # Count the total amount of simulations

			# Test if we have enough species at the timestep 150 in the simulation
  		success <- sum(simulation$pres[150,])
			#print(paste0("Success = ", success))

  		if(success >= 20){
				#print("the simulation has more than 19 species at the time step 150")
    		seed_record[successful_sim] <- seed # Record seed which alows us to have "good" simulation
    		res_sim <- list(simulation_number = paste0("simulation", successful_sim), 
    		                seed = seed,
    		                parameters = pars, 
    		                presence_matrix = simulation$pres, 
    		                traits_df = simulation$traits,
    		                parentage_matrix = simulation$anc,
    		                extinxction_matrix = simulation$extinct, 
    		                network_list = simulation$L_list, 
    		                anc_dist_table = simulation$dist_anc, 
    		                list_anc_dist = simulation$list_dist_anc) # Record results from the simulation
    		
    		if(int == 1){
    		  list_results_fac[[nsim]] <- res_sim
    		  
    		}else if(int == 0) {
    		  list_results_comp[[nsim]] <- res_sim
    		  
    		}
    		
    		
    		

    		print(paste("simulation ", successful_sim, "sur", nsim))
    		successful_sim <- successful_sim + 1 # Count the amount of "good" simulations
    		
    		
    		
    		
  		}
		}

		
	}


```


Pick simulation to inspect

```{r}
res <- list_results_fac[[2]]
res <- list_results_comp[[2]]
```







# Explore results


```{r}

# count number of timesteps where there are spp

n_steps <- length(res$network_list)

pres = res$presence_matrix



```



## N values with time

```{r}



res_spp = data.frame(matrix(ncol=2,
                      nrow=0, 
                      dimnames=list(NULL, c("step", "n values")))) #makes an empty dataframe

res_mean = data.frame(matrix(ncol=2,
                      nrow=0, 
                      dimnames=list(NULL, c("step", "n values")))) 


for (i in 1:nsteps) {
  
  pres_stepi <- which(pres[i,] == 1)
  traits_stepi <- res$traits_df[pres_stepi, "n"]
  
  vec_step <- rep(i, times = length(traits_stepi))
  
  
  
  res.add <- data.frame("step" = vec_step, 
                               "n values" = traits_stepi)
  
  res_spp <- rbind(res_spp, res.add)
  

  
  res.add1 <- data.frame("step" = i, 
                                "n values" = mean(traits_stepi))
  
  res_mean <- rbind(res_mean, res.add1)
  
  
}

plot_nichetime <- ggarrange(
  
  ggplot(res_spp, aes(x = step, y = n.values)) +
    geom_point(alpha = 0.2)+
    ggtitle("spp n values")+
    theme_classic()+
    my.theme+
    ylim(0,1),
  
  
  
  ggplot(res_mean, aes(x = step, y = n.values)) +
    geom_point(alpha = 0.2)+
    ggtitle("mean n values")+
    theme_classic()+
    my.theme+
    ylim(0,1),
  
  nrow = 2, 
  ncol = 1
  
  
)


plot_nichetime

```



## final degree distribution

```{r}

net_f <- res$network_list[[length(res$network_list)]]


#net_f <- net_f[26:1025,]

colnames(net_f) <- seq(1:1000)

rownames(net_f) <- seq(1:1000)

pres_f <- res$presence_matrix[length(res$network_list),]
names(pres_f) <- seq(1:1000)


net_f <- net_f[names(which(pres_f == 1)), names(which(pres_f == 1))]

graph <- graph_from_adjacency_matrix(adjmatrix = net_f,
                                     mode = "directed")

degree <- degree(graph, mode="all")

deg.dist <- degree_distribution(graph, cumulative=T, mode="all")

spp <- 1:length(deg.dist)

df_degreeplot <- data.frame(deg.dist, spp)

plot_degreedistr <- ggplot(df_degreeplot, aes(x = spp, y = deg.dist))+
  geom_point() +
  theme_classic()+
  my.theme+
  ylab("degree")+
  xlab("spp")

plot_degreedistr

```




## Speces richness with time



```{r}

plot_spp_richness <- function(steps, pres_mat, title) {
  
  timesteps <- 1:steps
  
  n_spp_time <- c()
  
  for (i in 1:nrow(pres_mat)) {
  
  n_spp_time[i] <- length(which(pres_mat[i,] == 1))
  
  }
  
  
  df_spprich_time <- data.frame(timesteps,n_spp_time)
  
  p.spp <- ggplot(df_spprich_time, aes(x=timesteps, y = n_spp_time)) +
  geom_line(color="black", linetype="twodash") +
  theme_classic()+
  my.theme+
  xlab("timesteps")+
  ylab("N species")+
    ggtitle(title)
  
  return(p.spp)
}


plot_richnesstime <- plot_spp_richness(steps = n_steps, pres_mat = res$presence_matrix[1:n_steps,], title = "foodweb")

plot_richnesstime
```


## Speciation and extinction events - time


```{r}

# Number of speciation and extinction events
t0 = pres[1:(n_steps-1),]
t1 = pres[2:n_steps,] 
spec_mat = pres[1:(n_steps-1),]*0
ext_mat = pres[1:(n_steps-1),]*0
spec_mat[t1-t0==1] = 1 
ext_mat[t1-t0==-1] = 1
spec = apply(spec_mat ,1,sum)
ext = apply(ext_mat, 1, sum)

S = apply(pres,1,sum)[2:n_steps]

time =  1:(n_steps-1)

df_divrates <- data.frame(spec, ext, S, time)

df_plot_divrates <- melt(df_divrates, id = c("time", "S"))


plot_divrates <- ggplot(data=df_plot_divrates, aes(x=time, y=value/S, group=variable)) +
  geom_line(aes(color = variable))+
  scale_color_manual(values = c("blue4", "red2"))+
  theme_classic()+
  my.theme+
  ylab("rate")+
  labs(color = "")

```



## Speciation and extinction events - S

```{r}
 ggplot(data=df_plot_divrates, aes(x=S, y=value/S, group=variable)) +
  geom_smooth(aes(color = variable), method = "loess") +
  scale_color_manual(values = c("blue4", "red2")) +
  theme_classic() +
  my.theme +
  ylab("rate") +
  labs(color = "")

```






## diversification-richness dependence

```{r}

plot_div_richness <- ggplot(data=df_plot_divrates, aes(x=S, y=value/S, group=variable)) +
  geom_point(aes(color = variable), size = 2, alpha = 0.4)+
  scale_color_manual(values = c("blue4", "red2"))+
  theme_classic()+
  my.theme+
  ylab("rate")+
  xlab("Species richness")+
  labs(color = "")

```



```{r}
arrange_plots <- ggarrange(
  plot_nichetime + ggtitle("evolution niche values"),
  plot_degreedistr + ggtitle("degree distr. final"),
  plot_richnesstime + ggtitle("evolution richness"),
  plot_divrates + ggtitle("evolution div. rates"),
  plot_div_richness +ggtitle("div.rates & richness")
)

arrange_plots



```
